---
title: "Continuous distributions"
author: "Fraser Tough"
date: "`r Sys.Date()`"
output:
  rmdformats::material:
    highlight: tango
    number_sections: true
    lightbox: true
    gallery: true
runtime: shiny
---

```{r knitr_init, echo=FALSE, cache=FALSE}
pziweibull=function(q,w,shape,scale)
{
  ifelse(q==0,w,(1-w)*pweibull(q,shape,scale)+w)
}

dziweibull=function(x,w,shape,scale)
{
  ifelse(x==0,w,dweibull(x,shape,scale)*(1-w))
}

qziweibull = function(p,w,shape,scale,br=c(0,10000))
{
  if(p<w){0}else{
    G = function(x) pziweibull(x,w,shape,scale)-p
    return( uniroot(G,br)$root ) 
  }
}
BoxCox=function(x,lambda)(x ^ lambda - 1)/lambda
InvBoxCox=function(x,lambda)(x*lambda +1)^(1/lambda)

suppressMessages(library(knitr))
suppressMessages(library(rmdformats))
suppressMessages(library(lubridate))
suppressMessages(library(plyr))
suppressMessages(library(MASS))
suppressMessages(library(gamlss))
suppressMessages(library(data.table))
suppressMessages(library(viridis))
## Global options
#options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=FALSE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
#opts_knit$set(width=75)
```

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# The problem with rainfall 

Before developers carry out development, they need to characterise and understand precipitation in the local region. The vast majority of the planet observes seasonal variation in rainfall which, at many times of year, can result in the inability to carry out works. If precipitation levels are poorly understood, leading to misinformed decision making, this can result to project delay or even property damage. In regions with large seasonal variation, understanding the likely levels of rainfall prior to construction can help planners manage when to start works, minimising the probability of delay.

Rainfall distributions, therefore, play a pivotal role in project planning and construction design.


## The distribution of rainfall
## Weibull mixture model

Rainfall distributions are bounded by zero, have long tails and can generally be well characterised by the 2 parameter Weibull. In locations with heavy average rainfall, the 3 parameter Weibull can also be used. However, the Weibull is undefined at 0 and the lower the temporal resolution, the higher the probability that zeros will occur. 

The shape and scale of the distributions varies by time of year. In northern Europe, for example, rainfall is heaviest during the winter. In summer, it is likely that daily average rainfall will be zero cm per day.

Characterising all three of these parameters, scale, shape and probability of zero requires a mixture model. The probability that rainfall will be positive if Bernoulli distributed and if rain does occur, the probability density is Weibull distributed. 

The following mixture model can be used,

\begin{align}
f(y) =
\begin{cases}
 p\left(\frac{k}{\lambda}\Bigl(\frac{x}{k}\Bigr)^{k-1} \exp\bigl(-(x/\lambda)^k\bigr)\right) & x > 0 \\[1ex]

 (1-p)  & x=0
 \end{cases}       
\end{align}

which consists of both discrete (Bernoulli) and continuous (Weibull) parts.


```{r, echo=FALSE}
shinyAppFile(
  "./apps/ZIW.R",
  options = list(
    width = "100%", height = 550
  )
)
```

## Modelling seasonailty via harmonics

Most environmental trends exhibit diurnal or seasonal trend which smoothly varies across time. These trends tend to 
Any periodic function can be represented as a sum of cosine waves and 
Seasonal trend is circular. Values of the variable of interest at start and endpoints of a year should match

Generally, seasonal trend can be well captured by


Fourier components are 
```{r, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)

#############################################
##################READ DATA IN###############
#############################################

Data=data.frame(fread("./data/killermont rainfall.csv",skip=20))[,1:2]
colnames(Data)<-c("DATE","PRCP")

Data$DATE=dmy(Data$DATE)
Data$Month=factor(month(Data$DATE))
Data$Year=year(Data$DATE)
Data$PRCP=ifelse(is.na(Data$PRCP),0,Data$PRCP)
Data$DecimalDate=decimal_date(Data$DATE)
Data$DecimalYear=decimal_date(Data$DATE)-year(Data$DATE)
Data$PRCP_cat=factor(ifelse(Data$PRCP>0,1,0))
Data$yday=yday(Data$DATE)

#############################################
##################Models###############
#############################################
newdata=data.frame("DecimalYear"=seq(0,1,length.out=365))

gamlss=gamlss(PRCP~sin(2*pi*DecimalYear)+cos(2*pi*DecimalYear),
              sigma.fo=~1,data=Data[Data$PRCP>0,c("PRCP","DecimalYear")],family="WEI",trace=FALSE)
glm=glm(PRCP_cat~sin(2*pi*DecimalYear)+cos(2*pi*DecimalYear),data=Data,family="binomial")

p_gamlss=predictAll(gamlss,newdata=newdata,data=Data[Data$PRCP>0,c("PRCP","DecimalYear")])
p_glm=predict(glm,newdata=newdata,type="response")

newdata$yday=1:365
newdata$w=p_glm
newdata$shape=p_gamlss$sigma
newdata$scale=p_gamlss$mu



#############################################
##################Quantiles###############
#############################################
Q=c(0,0.95,0.96,0.97,0.99,0.995,0.996,0.997,0.998,0.999)
col=viridis(length(Q),alpha=0.5)

quantiles=data.frame(t(sapply(1:nrow(newdata),function(i){sapply(Q,function(s)qziweibull(s,w=newdata$w[i],shape=newdata$shape[i],scale=newdata$scale[i]))})))
quantile_names=paste0("Q",Q*100)
colnames(quantiles)<-quantile_names
quantiles=cbind(newdata,quantiles)

# #setwd(paste0(getwd(),"/data"))
# data_flow=data.frame(fread("./data/killermont flow.csv",skip=20))[,1:2]
# colnames(data_flow)<-c("datetime","flow")
# data_flow$datetime=dmy(data_flow$datetime)
# 
# data_list_rainfall=lapply(list.files(recursive = TRUE)[grepl("rainfall",list.files(recursive = TRUE))],function(x){
#   data=data.frame(fread(x,skip=20))[,1:2]
#   colnames(data)<-c("datetime",paste0("rainfall_",substr(x,1,regexpr(" ",x)-1)))
# 
#   if(is.na(ymd(data$datetime[1]))){
#   data$datetime=dmy(data$datetime)
#   }else{
#     data$datetime=ymd(data$datetime)
#   }
#   return(data)
#   })
# names(data_list_rainfall)<-sapply(list.files()[grepl("rainfall",list.files())],function(x)substr(x,1,regexpr(" ",x)-1))
# data_rainfall=join_all(data_list_rainfall,by="datetime")
# data=join(data_flow,data_rainfall,by="datetime")
# data=data[complete.cases(data),]
# head(data)
# #BX=boxcox(lm(flow~1,data),plotit = FALSE)
# 
# #Set Box-Cox
# # BX=boxcox(lm(flow~1,data))
# # lambda=BX$x[which.max(BX$y)]+0.001
# # lambda= -0.199
# # data$tflow=BoxCox(data$flow,lambda)
# data$DecimalDate=decimal_date(data$datetime)
# data$DecimalYear=decimal_date(data$datetime)-year(data$datetime)
# data=data[complete.cases(data),]

```

\begin{align}
E(y|x)=\alpha+\beta cos(2\pi-\gamma)
\end{align}

\begin{align}
E(y|x)=\alpha+A sin(2\pi-\gamma)+B sin(2\pi-\gamma)
\newline
\alpha=\alpha
\newline
\beta=\sqrt{A^2+B^2}
\newline
\gamma=atan2(A,B)

\end{align}



```{r, echo=FALSE}
shinyAppFile(
  "./apps/harmonics.R",
  options = list(
    width = "80%", height = 550
  )
)
```


Carry out construction at a different time of day or year
hourly uncertainty - this is just the average of the day
Rainfall vs. flow level ( download flow data further up river)
different distributions - compare AIC using adaption - prove that AIC is same as leave one out cross



```{r, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)

#Plots
 par(bg="black",fg="white",col.lab="white",col.main="white",col.lab="white",col.axis="white",col.sub="white")
plot(Data$PRCP~Data$yday,pch=20,cex=0.1,ylab="Precipitation (in.)",xlab="Day of year")
    abline(v=seq(0,max(Data$yday),length.out = 11),col="grey20",lty=2)
    abline(h=seq(0,max(Data$PRCP),length.out = 11),col="grey20",lty=2)

sapply(length(quantile_names):2,function(i){
  suppressMessages(polygon(c(quantiles$yday,rev(quantiles$yday)),c(quantiles[[quantile_names[i]]],rev(quantiles[[quantile_names[i-1]]])),col=col[i],border=NA))
})
legend("topright",legend=paste0("Centiles ",rev(c("10-20","20-30","30-40","40-50","50-60","60-70","80-90"))),col=rev(col),lwd=10,bty="n")
```

